Ethnographic Study of Online Entertainers

### 2017-03-16 proposals/Brown_Magic_Grant_2017/main.tex
framing rough draft done?

\begin{document}
  \maketitle
  \begin{abstract}
  TODO? ahhhh
  \end{abstract}

\section*{Situation}
\topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
and streaming sites % \ali{, like Twitch and others,}
have garnered the attention of audiences around the world
in the span of little more than a decade.} % taken on characteristics of
So widely used are these platforms
that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
content producers
--- ``YouTubers'' and ``streamers'' ---
have begun to make their livings on these sites.
Indeed, these performers have come to rely on
the technical infrastructure these platforms provide as well as
the access to audiences that these platforms provide as well.
In doing so, these sites have in a real sense democratized media and entertainment industries by
making the avenues to large audiences substantially more accessible.
And thus, in these transformations into publics, what started as
a recreational hobby
has become a bustling hub for professional artists to
engage with audiences,
practice their craft, and 
hone their skills~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}.

\section*{Problem}
\topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
assumptions regarding the nature of the space they're using.}
Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
\ugh{These events fundamentally speak to the contentious politics of an emerging form of expression and art,
and specifically to the uncertainty regarding the status of these artists who tend to subvert not just
the norms of their chosen arts, but also
orthodox labels of performers in the first place.}

\section*{Insight}
\topic{Studying contemporary phenomena through an established framing
can be a helpful way of
making sense of what we've seen so far, as well as
providing us with some grounding to make predictions.}
We've done this before, studying crowd work and gig work as
a modern instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork}.
This positioning of an existing subject in historical context not only
provided us with guidance toward some of the major threads of research
in crowdsourcing, but
also informed some of the potential futures of crowdsourcing and
the ever growing gig work labor markets.

@@

\begin{blah}
Online performers
tend not to have the formal training of many conventional performers and artists,
instead growing in proficiency and becoming more skilled principally through their practice;
in much the same way,
street performers are 
%
Online performance artists have the ability
to engage directly with their audiences and tailor their performances to the desires of the crowd
--- YouTubers call upon their audiences to like, dislike, and submit comments to the video, while
    streamers literally change their performances on the fly to suit the whims of those watching and chatting;
meanwhile, the most successful street performers typically engage participants,
drawing them into the performance itself, even bantering with passersby.
%
Finally, online performance has increasingly become a site of political engagement and activism,
sometimes subverting authority
--- in these cases online moderators and algorithms ---
in the process of or for the purpose of making a point;
the history of street performance as a channel for political action and advocacy has been deeply studied
(see, for example, \cite{harrison1990drawing,cohen1998radical}).
\end{blah}

\section*{Incomplete!}
\section*{Proposal}

% \section*{Motivating Work: Historical Analysis of Contemporary Phenomena}

% Online performers share an uncanny number of similarities with street performers

% \ali{generally speaking, each major bullet point = roughly 1 paragraph}
\begin{itemize}
  \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
        a growing industry for entertainers.
        Further, it has ``democratized'' entertainment, providing people with a public space
        (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
  \item Our research into crowdsourcing and gig work has yielded
        a compelling framework for these phenomena as an instantiation of piecework.
        We've found that this approach
        --- that is, drawing parallels with historical phenomena ---
        can generally give researchers of sociotechnical systems substantive framing of a research space.
  \item To some extent, this can be likened to \textbf{street performance} inasmuch as
  \begin{itemize}
    \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
    \item entertainers could get more direct feedback from their audience than other avenues could afford them
    \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
  \end{itemize}
  \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
        \ali{need a clearer \textbf{research question}?}
  \item When we look at this dimension of online performance art, the metaphor begins to break down:
  \begin{itemize}
    \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
    \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
    \item communication between the performer and the audience is carefully mediated
          (and literally moderated) by the platform itself
    \item online performances are (generally) more permanent than street performances
          \ali{technology has thrown a wrench into this lately, with
              cameras allowing us to record street performances and
              tools making it impossible (or at least difficult) to save online streams, but
              \textit{let's not worry about all that}.}
  \end{itemize}
  \item So while there are compelling parallels between online performances and the street art of busking,
        this has only taken us partway in answering important questions about a rapidly growing field.
\end{itemize}

\section*{Research Plan}
% \begin{itemize}
%   \item We propose to direct our research energy toward answering one question: 
% \end{itemize}

>>>>

\begin{document}
  \maketitle
  \begin{abstract}
  \ali{something along the lines of: online performers are becoming a big deal, and
       these performers are possibly the forerunners of a new type of expression and art.
       The frustration and turmoil they're experiencing
       --- both interpersonally and with the environment on which they ``work'' ---
       demands to be better understood if we hope to make informed design decisions that empower these people.
       We propose to conduct ethnographic fieldwork of this domain,
       initially informed by a historical framing, but
       ultimately studying how digitally mediated publics such as these affect
       the politics of this group of performers.}
  \end{abstract}

% \section*{Situation}
\topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
and streaming sites % \ali{, like Twitch and others,}
have garnered the attention of audiences around the world
in the span of little more than a decade.} % taken on characteristics of
So widely used are these platforms
that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
content producers
--- ``YouTubers'' and ``streamers'' ---
have begun to make their livings on these sites.
Indeed, these performers have come to rely on
the technical infrastructure these platforms provide as well as
the access to audiences that these platforms provide as well.
In doing so, these sites have in a real sense democratized media and entertainment industries by
making the avenues to large audiences substantially more accessible.
And thus, in these transformations into publics, what started as
a recreational hobby
has become a bustling hub for professional artists to
engage with audiences,
practice their craft, and 
hone their skills~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}.

% \section*{Problem}
\topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
assumptions regarding the nature of the space they're using.}
Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
These events highlight the politics of an emerging form of expression and art,
and specifically the sometimes controversial nature of the experimentation that performers engage in.

% \section*{Insight}
\topic{Studying contemporary phenomena through an established framing
can be a helpful way of
making sense of what we've seen so far, as well as
providing us with some grounding to make predictions.}
We've done this before, studying crowd work and gig work as
a modern instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork}.
This positioning of an existing subject in historical context not only
provided us with guidance toward some of the major threads of research
in crowdsourcing, but
also informed some of the potential futures of crowdsourcing and
the ever growing gig work labor markets.

@@

\topic{This connection has purchase, but it has flaws, mostly owing to the unique nature of ``networked publics''~\cite{boyd2007youth}.}
For one thing, most video sharing websites keep videos
% --- the modern performances ---
in perpetuity
--- that is, unless a complaint over copyright infringement is filed, in which case the lifespan of that ``performance'' can be cut short.
The public that congregates around a performer, too, has similar permanence that offline busking doesn't;
YouTube comments linger for years after they're made.

% \section*{TODO: Problem with this model}
% \section*{TODO: Proposal}
% \section*{Incomplete!}
% \section*{Proposal}

% \ali{generally speaking, each major bullet point = roughly 1 paragraph}
% \begin{itemize}
%   \item Online streaming, which started as a recreational hobby~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}, has become
%         a growing industry for entertainers.
%         Further, it has ``democratized'' entertainment, providing people with a public space
%         (e.g. YouTube, Twitch, Vimeo, etc\dots) to explore unconventional performance art
%   \item Our research into crowdsourcing and gig work has yielded
%         a compelling framework for these phenomena as an instantiation of piecework.
%         We've found that this approach
%         --- that is, drawing parallels with historical phenomena ---
%         can generally give researchers of sociotechnical systems substantive framing of a research space.
%   \item To some extent, this can be likened to \textbf{street performance} inasmuch as
%   \begin{itemize}
%     \item the public space allowed non--experts to engage in their craft without commitment, formal training, etc\dots
%     \item entertainers could get more direct feedback from their audience than other avenues could afford them
%     \item entertainers could (and did) subvert authorities (offline, that was the police; online, that's YouTube's moderator team)
%   \end{itemize}
%   \item \textbf{But} there are questions about the politics of (grassroots, potentially subversive) performance art.
%         \ali{need a clearer \textbf{research question}?}
%   \item When we look at this dimension of online performance art, the metaphor begins to break down:
%   \begin{itemize}
%     \item policies and rules are enacted and enforced opaquely, seemingly arbitrarily on private settings
%     \item the notion of the public in this instance is broken, as the phenomenon of shared experience doesn't translate online
%     \item communication between the performer and the audience is carefully mediated
%           (and literally moderated) by the platform itself
%     \item online performances are (generally) more permanent than street performances
%           \ali{technology has thrown a wrench into this lately, with
%               cameras allowing us to record street performances and
%               tools making it impossible (or at least difficult) to save online streams, but
%               \textit{let's not worry about all that}.}
%   \end{itemize}
%   \item So while there are compelling parallels between online performances and the street art of busking,
%         this has only taken us partway in answering important questions about a rapidly growing field.
% \end{itemize}

\section*{Research Proposal}

\section*{Broader Impact}
\topic{Formally recognized art forms have assimilated emergent genres and types of performance art in the past, and
we should expect that they will continue to do so.}
Jazz was born out of what largely ``\dots was recognized~\dots~as being but one species of the genus, ragtime played by ear by fakers\dots''~\cite{10.2307/779456}.
Even further back in history, \nth{17} century street performance in Japan blended ``\dots every imaginable genus and species of popular song and dance\dots'' and
kabuki, in its early years,
``would hardly have been recognized or condoned [by traditional performers]''~\cite{groemer2016street}.

\ali{something like: if this continues, then we should infer that
     the future of this type of performance art will bear heavily on
     the direction and trajectory of digitally mediated entertainment and performance}

\section*{Demonstrate viability}

\section*{Expected outcomes}

% \begin{itemize}
%   \item We propose to direct our research energy toward answering one question: 
% \end{itemize}
